---
title: "Task 8 Functions"
output:
  md_document:
    variant: markdown_github
---

#Write a Function

* Complete this task by creating an R notebook (.Rmd file) and accompanying .md file ## Functions and error handling: Write a function that fulfills the following criteria:

  * It should be tidyverse compatible (i.e., the first argument must be a data frame)
  
  * It should add two arbitrary columns of the data frame (specified by the user) and put them in a new column of that data frame, with the name of the new column specified by the user
  
```{r}
# load libraries
library(tidyverse)
library(ggplot2)

column_addition <- function(df, col1, col2, name = "summation"){
  if (class(df) != "data.frame"){
    stop("Your first argument is not a data frame!")
  }
  
  if (class(col1) != "numeric" || class(col2) != "numeric"){
    #warning("your columns are non-numeric")
    #why does the function change the col1 and col2 to character
  }

  df[[name]] = df[[col1]] + df[[col2]]
  df
}

#source('~/University/Classes/590_R/R/task_8_function_2.R')
```


```{r}
#test the function
column_addition(mtcars, "mpg", "cyl", name = "add")

```

  * It should throw an informative warning if any invalid arguments are provided. Invalid arguments might include:
  
    * The first argument is not a data frame
    
    * Less than two valid columns are specified to add (e.g., one or both of the column names isn't in the supplied data frame)
    
    * The columns specified are not numeric, and therefore can't be added - use tryCatch() for this

  * If the columns to add aren't valid but the new column name is, the function should create a column of NA values
  
#Loop and performance metric task

* Write a function named that uses a for loop to calculate the sum of the elements of a vector, which is passed as an argument (i.e., it should do the same thing that sum() does with vectors). your_fun(1:10^4) should return 50005000.
```{r}
#### write a for loop
my_sum <- function(x){
  r <- c(x)
  s <- 0
  
  for (i in r) {
    s <- s + i
  }
  s
}
```

```{r}
# test my loop
my_sum(1:10^4)
```

```{r}
# test the sum function
sum(1:10^4)
```




* Use the microbenchmark::microbenchmark function to compare the performace of your function to that of sum in adding up the elements of the vector 1:10^4. The benchmarking code should look something like:

    test.vec <- 1:10^4
microbenchmark(
    my_sum(test.vec),
    sum(test.vec)
    )

```{r}
#use the microbenchmark function to compare performance
library(microbenchmark)
test.vec <- 1:10^4
microbenchmark(
  my_sum(test.vec),
  sum(test.vec)
  )
```

    
Is there a difference? Why?
* the for loop takes by far longer to run